---
title: "Exercise 1"
author: "Ryan Clorfeine"
date: "2/5/2020"
output: 
  md_document:
    variant: markdown_github
---

# Exercise 1
```{r message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(mosaic)
urlfile = "https://raw.githubusercontent.com/jgscott/SDS323/master/data/ABIA.csv"
mydata <- read_csv(url(urlfile))
```

## Data visualization: flights at ABIA
```{r}
# Group data by airline and calculate mean arrival/departure delay
grouped_Data <- mydata %>% group_by(UniqueCarrier) %>% summarise(mean_DepDelay=mean(DepDelay,na.rm=TRUE),mean_ArrDelay=mean(ArrDelay,na.rm=TRUE), count=n())

#Get subset of top 5 airlines in terms of number of flights.
grouped_Data <- grouped_Data %>% arrange(desc(count))
grouped_Data <- grouped_Data %>% slice(1:5)

# Flatten data to seperate arrival and departure delay
grouped_Data <- grouped_Data %>% gather("Direction","Delay",-count,-UniqueCarrier)

grouped_Data$UniqueCarrier <- recode(grouped_Data$UniqueCarrier, AA="American", WN="Southwest",B6="JetBlue", YV="Mesa", CO="Continental")
ggplot(grouped_Data,aes(x=UniqueCarrier,y=Delay,fill=Direction)) +
  geom_col(position = "dodge") +
  ggtitle("Austin Flight Arrival/Departure Delay by Airline")+
  labs(x="Airline",y="Mean Delay")

```

<ins>Conclusion</ins>
<br>
From the graph, it is reasonable to conclude that their is not any significant difference in flight delays among the top 5 major airlines flying out of Austin. The exception is Southwest Airlines which has the lowest arrival delay. The reason for this could be that many of Southwest flight into Austin are in-state flights and thus have less time for delays to occur.

## Regression Practice

```{r}
creatinine = read.csv("creatinine.csv",header=TRUE)
lm1 = lm(creatclear~age,data = creatinine)

#plug in age 55 to linear model to get expected creatine clearance rate
as.numeric(coef(lm1)[1]) + (55 * as.numeric(coef(lm1)[2])) 
```
1. Based on the linear model we can expect a creatine clearance rate of 113.723 ml/minute
